### 2023-05-01 thesis/chapters/02-analysis-A-CdM.tex
Describe MaFaulDa dataset

One of the methods vibration experts utilize in the identification of the damaged part from the frequency spectrum is \textbf{order analysis}. In essence, it is about noticing the excessive peaks at harmonic frequencies that are integer multiples of fundamental frequency (1x RPM)~(Tab.~\ref{tab:vibration-causes}):

\begin{table}[h]
\renewcommand{\arraystretch}{1.2}
\begin{adjustbox}{width=\columnwidth,center}
\begin{tabular}{|ll|l|l|}
\hline
\multicolumn{2}{|l|}{\textbf{Frequency content}}                            & \textbf{Likely reason}                                                                                                     & \textbf{Other causes}                                                                                                                          \\ \hline
\multicolumn{1}{|l|}{\multirow{4}{*}{Synchronous}} & 1 x RPM                & Imbalance                                                                                                                & \begin{tabular}[c]{@{}l@{}}Eccentric journals\\ Bent shaft / Misalignment\\    (high axial vibration)\\ Bad belt (if RPM of belt)\end{tabular} \\ \cline{2-4}
\multicolumn{1}{|l|}{}                             & 2 x RPM                & Looseness                                                                                                                & \begin{tabular}[c]{@{}l@{}}Misalignment \\    (high axial vibration)\\ Cracked rotor\\ Bad belt (if RPM of belt)\end{tabular}                  \\ \cline{2-4}
\multicolumn{1}{|l|}{}                             & 3 x RPM                & Misalignment                                                                                                             & and axial looseness                                                                                                                             \\ \cline{2-4}
\multicolumn{1}{|l|}{}                             & Many x RPM             & \begin{tabular}[c]{@{}l@{}}Bad gears\\ Severe looseness\end{tabular}                                                     & \begin{tabular}[c]{@{}l@{}}Gear teeth x RPM\\ Fan blade count x RPM\end{tabular}                                                               \\ \hline
\multicolumn{1}{|l|}{Sub-synchronous}                 & \textless 1 x RPM      & Oil whirl                                                                                                                & \begin{tabular}[c]{@{}l@{}}Bad drive belt\\ Background\\ Resonance\end{tabular}                                                                \\ \hline
\multicolumn{1}{|l|}{Non-synchronous}              & \multicolumn{1}{c|}{-} & \begin{tabular}[c]{@{}l@{}}Electrical problems (x 50 Hz)\\ Reciprocating forces\\ Aerodynamic forces \\ Bad antifriction bearings\end{tabular} & Rubbing \\ \hline
\end{tabular}
\end{adjustbox}
\caption{Expert observed likely vibration causes (based on~\cite{davies_handbook_2012,ziaran_technicka_2013,noauthor_iso_2002})}
\label{tab:vibration-causes}
\end{table}

>>>>

One of the methods vibration experts utilize in the identification of the damaged part from the frequency spectrum is \textbf{order analysis}. In essence, it is about noticing the excessive peaks at harmonic frequencies that are integer multiples of fundamental frequency (1x rpm)~(Tab.~\ref{tab:vibration-causes}):

\begin{table}[h]
\renewcommand{\arraystretch}{1.2}
\begin{adjustbox}{width=\columnwidth,center}
\begin{tabular}{|ll|l|l|}
\hline
\multicolumn{2}{|l|}{\textbf{Frequency content}}                            & \textbf{Likely reason}                                                                                                     & \textbf{Other causes}                                                                                                                          \\ \hline
\multicolumn{1}{|l|}{\multirow{4}{*}{Synchronous}} & 1 x rpm                & Imbalance                                                                                                                & \begin{tabular}[c]{@{}l@{}}Eccentric journals\\ Bent shaft / Misalignment\\    (high axial vibration)\\ Bad belt (if rpm of belt)\end{tabular} \\ \cline{2-4}
\multicolumn{1}{|l|}{}                             & 2 x rpm                & Looseness                                                                                                                & \begin{tabular}[c]{@{}l@{}}Misalignment \\    (high axial vibration)\\ Cracked rotor\\ Bad belt (if rpm of belt)\end{tabular}                  \\ \cline{2-4}
\multicolumn{1}{|l|}{}                             & 3 x rpm                & Misalignment                                                                                                             & and axial looseness                                                                                                                             \\ \cline{2-4}
\multicolumn{1}{|l|}{}                             & Many x rpm             & \begin{tabular}[c]{@{}l@{}}Bad gears\\ Severe looseness\end{tabular}                                                     & \begin{tabular}[c]{@{}l@{}}Gear teeth x rpm\\ Fan blade count x rpm\end{tabular}                                                               \\ \hline
\multicolumn{1}{|l|}{Sub-synchronous}                 & \textless 1 x rpm      & Oil whirl                                                                                                                & \begin{tabular}[c]{@{}l@{}}Bad drive belt\\ Background\\ Resonance\end{tabular}                                                                \\ \hline
\multicolumn{1}{|l|}{Non-synchronous}              & \multicolumn{1}{c|}{-} & \begin{tabular}[c]{@{}l@{}}Electrical problems (x 50 Hz)\\ Reciprocating forces\\ Aerodynamic forces \\ Bad antifriction bearings\end{tabular} & Rubbing \\ \hline
\end{tabular}
\end{adjustbox}
\caption{Expert observed likely vibration causes (based on~\cite{davies_handbook_2012,ziaran_technicka_2013,noauthor_iso_2002})}
\label{tab:vibration-causes}
\end{table}

### 2023-05-01 thesis/chapters/02-analysis-D-datasets.tex
Describe MaFaulDa dataset

\section{Evaluation datasets}

To validate the industrial solution beforehand on comprehensive dataset we use following as a benchmark.  Faults in the wild are rare so the preparation of balanced datasets is done create deficient configuartions on purpose. Datsets MaFaulDa and CWRU bearing dataset are being used in related work and are publically available in the Comma-Separated Values format. There is also less comprehensive dataset conserning unbalance, but demostrates behaviour during revolution speed up.

\subsection{Machinery Fault Database}
(MaFaulDa)
50 kHz, 5 sec. recordings, 
Rotation speeds 10~Hz - 60~Hz (737 rpm - 3686 rpm with steps of approximately 60 rpm)

% Sensors: 
	% - Accelerometers IMI 601A01/604B31,  Range ±50 g (±490 m/s2)
	% Magnetic tachometer MT-190
	% Shure SM81 microphone with frequency range of 20-20.000 Hz
	% National Instruments NI 9234 4 channel analog acquisition modules, with sample rate of 51.2 kHz

% Normal
% Imbalance  - load values within the range from 6 g to 35 g (step of 5g)
% Misalignment (Horizontal/Vertical)
% Bearings (Overhang / Underhang) - Inner, Outer, Cage

% total of 1951 different fault scenarios for 4 different operational conditions. 
% 49 of which from the normal class, 333 from unbalance class, 498 from misalignment class and 1071 from the bearing fault.

% 2 sets of accelerometers each one associated to one bearing (inner and outer) and measuring in 3 directions

% Column 1 - tachometer signal that allows to estimate rotation frequency
% Columns 2 to 4 - underhang bearing accelerometer (axial, radiale tangential direction)
% Columns 5 to 7 - overhang bearing accelerometer (axial, radiale tangential direction)
% Column 8 - microphone

RotorKit Alignment Balance Vibration Trainer (ABVT)
\cite{pestana-viana_influence_2016}
\cite{ribeiro_rotating_2017}

 \footnote{\url{https://www02.smt.ufrj.br/~offshore/mfs/page_01.html}}

\begin{figure}[h]
\centering
\begin{subfigure}[b]{0.48\textwidth}
	\includegraphics[width=\textwidth]{assets/mafaulda-simulator.png}
	\caption{Schematic diagram \cite{pestana-viana_influence_2016}}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.48\textwidth}
	\includegraphics[width=\textwidth]{assets/machinery-fault-simulator.jpg}
	\caption{Mechanical construction \cite{noauthor_spectraquest_nodate}}
\end{subfigure}
\caption{Machinery Fault Simulator for MaFaulDa}
\label{fig:mafaulda-simulator}
\end{figure}

\subsection{Case Western Reserve University bearing database}
(CWRU) 
2 HP (1.492 kW) Reliance Electric motor \footnote{\url{https://engineering.case.edu/bearingdatacenter/download-data-file}}
Bearings - Inner, Outer
12 kHz, 48 kHz
fan and drive end bearings
Fault diameters of 7 mils, 14 mils, 21 mils, 28 mils, and 40 mils (1 mil=0.001 inches) in diameter were introduced separately at the inner raceway, rolling element (i.e. ball) and outer raceway.

@@

\begin{figure}[h]
\centering
\begin{subfigure}[b]{0.48\textwidth}
	\includegraphics[width=\textwidth]{assets/cwru-test-stand-2.png}
	\caption{Schematic diagram \cite{song_bearing_2022}}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.48\textwidth}
	\includegraphics[width=\textwidth]{assets/cwru-test-stand.png}
	\caption{Mechanical construction \cite{yuhong_new_2021}}
\end{subfigure}
\caption{CWRU apparatus}
\label{fig:cwru-simulator}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.7\textwidth]{assets/rotating-shaft.jpg}
\caption{Rotating shaft dataset \cite{mey_machine_2020}}
\label{fig:rotating-shaft}
\end{figure}

>>>>

\section{Evaluation datasets}
Before the experimental setting testing designed features relevancy is proven in comparison to comprehensive benchmark datasets. There are a few standardized datasets used in the related work \cite{ribeiro_rotating_2017} and are publically available in Comma-Separated Values files.

\emph{MaFaulDa} dataset combines vibration and acoustic measurements of the shaft in deviating positions and bearings abnormalities.\emph{CWRU dataset} focuses solely on faults in ball bearings. Another less known dataset concerns shaft unbalance, but compared to the previous two, it demonstrates behavior during revolution speed up.

\subsection{Machinery Fault Database}
MaFaulDa\footnote{MaFaulDa: \url{https://www02.smt.ufrj.br/~offshore/mfs/page_01.html}} is a collection of 1951 multivariate time series for 4 different operational conditions on rotor kit Alignment Balance Vibration Trainer (ABVT)~(Fig.~\ref{fig:mafaulda-simulator}). Each series has 5~seconds in length and is captured at 50 kHz. Vibration signals were obtained with piezoelectric accelerometers with a linear response up to 10~kHz, amplitude range to 490 $m/s^2$, and resolution step of 10.2~mV per $m/s^2$.

\begin{figure}[h]
\centering
\begin{subfigure}[b]{0.48\textwidth}
	\includegraphics[width=\textwidth]{assets/mafaulda-simulator.png}
	\caption{Schematic diagram \cite{pestana-viana_influence_2016}}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.48\textwidth}
	\includegraphics[width=\textwidth]{assets/machinery-fault-simulator.jpg}
	\caption{Mechanical construction \cite{noauthor_spectraquest_nodate}}
\end{subfigure}
\caption{Machinery fault simulator for MaFaulDa}
\label{fig:mafaulda-simulator}
\end{figure}

Observations were conducted in three cardinal axes simultaneously with 2 sets of accelerometers each one associated with one bearing (inner and outer bearings)~(Fig.~\ref{fig:mafaulda-simulator}). Additionally, a magnetic tachometer produced a pulse on shaft turn. The cardioid condenser microphone recorded sound emissions with frequency range 20~Hz - 20~kHz. Sensors were fed into a four-channel dynamic signal acquisition module.

Columns in dataset are organized as depicted in table~\ref{tab:mafaulda-columns}. Machine rotational speeds were kept constant during a particular measurement, but covered a range from 737 to 3686~rpm with steps of approximately 60 rpm (equiv. 10~Hz - 60~Hz)~\cite{pestana-viana_influence_2016}. The maximal rotational frequency achieved with high unbalance load is 3300 rpm.

\begin{table}[h]
\renewcommand{\arraystretch}{1.2}
\centering
\begin{tabular}{|l|l|}
\hline
\textbf{Columns} & \textbf{Description}                                                                                                                                               \\ \hline
1.               & \begin{tabular}[c]{@{}l@{}}Pulse with modulation of tachometer signal \\ to estimate rotation frequency  (in TTL levels)\end{tabular}                              \\ \hline
2., 3., 4.       & \begin{tabular}[c]{@{}l@{}}Underhang bearing accelerometer \\ (inner - between the rotor and motor)\\ - axial, radial, tangential direction\end{tabular}           \\ \hline
5., 6., 7.       & \begin{tabular}[c]{@{}l@{}}Overhang  bearing accelerometer \\ (Outer - outside most position after the rotor)\\ - axial, radial, tangential direction\end{tabular} \\ \hline
8.               & Microphone                                                                                                                                                         \\ \hline
\end{tabular}
\caption{MaFaulDa columns' descriptions}
\label{tab:mafaulda-columns}
\end{table}

This database contains normal operating conditions, faults out of unbalance, horizontal and vertical shaft misalignment, and three types of faulty bearings in inner and outer positions: outer track, inner track, rolling elements~\cite{pestana-viana_influence_2016}.
\begin{itemize}
\itemsep0pt
\item \textbf{Normal} conditions are baseline without the adverse effect of fault in 49 different rotation speeds. 
\item \textbf{Unbalance} shaft time series uses 8 unbalancing weights from 6 to 35 grams and varying 45 - 49 speeds for each weight adding to 333 mass unbalance loads. 
\item \textbf{Vertical misalignement} set is comprised of 50 signals each (or 51 in one instance) obtained under displacements: 0.51, 0.63, 1.40, 1.90, 1.27, 1.78 mm.
\item \textbf{Horizontal misalignement} signals were recorded under displacements: 0.50, 1.00, 1.50, 2.00 mm, each with 49 different speeds (or 50 in one instance)~\cite{pestana-viana_influence_2016}.
\item \textbf{Bearing faults} are unnoticable without unbalance. Therefore, weights of 6, 20, and 35 grams were attached to induce a detectable effect. Each unbalance mass was combined with cage, outer race, and ball faults at multiple rotation speeds, usually at 50 different speeds.
\end{itemize}

%TODO
\subsection{Case Western Reserve University bearing database}
(CWRU) 
2 HP (1.492 kW) Reliance Electric motor \footnote{\url{https://engineering.case.edu/bearingdatacenter/download-data-file}}
Bearings - Inner, Outer
12 kHz, 48 kHz
fan and drive end bearings
Fault diameters of 7 mils, 14 mils, 21 mils, 28 mils, and 40 mils (1 mil=0.001 inches) in diameter were introduced separately at the inner raceway, rolling element (i.e. ball) and outer raceway.

@@

\begin{figure}[h]
\centering
\begin{subfigure}[b]{0.48\textwidth}
	\includegraphics[width=\textwidth]{assets/cwru-test-stand-2.png}
	\caption{Schematic diagram \cite{song_bearing_2022}}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.48\textwidth}
	\includegraphics[width=\textwidth]{assets/cwru-test-stand.png}
	\caption{Mechanical construction \cite{yuhong_new_2021}}
\end{subfigure}
\caption{CWRU machine apparatus}
\label{fig:cwru-simulator}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.7\textwidth]{assets/rotating-shaft.jpg}
\caption{Motor driving shaft in unbalance measurement \cite{mey_machine_2020}}
\label{fig:rotating-shaft}
\end{figure}
